\documentclass[11pt]{article}

\author{Math 123}
\date{Due April 7, 2023 by midnight} 
\title{Final Project Proposal}

\usepackage{graphicx,xypic}
\usepackage{amsthm}
\usepackage{amsmath,amssymb}
\usepackage{amsfonts}
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage[shortlabels]{enumitem}
\newtheorem{problem}{Problem}
\renewcommand*{\proofname}{{\color{blue}Solution}}


\usepackage{fancyhdr}
\pagestyle{fancy}
\rhead{Math 123, Final project proposal}

\setlength{\parindent}{0pt}
\setlength{\parskip}{1.25ex}

\input{../tikz_import.tex}

\begin{document}

\maketitle

% You are required to put your name here:
{\bf\Large Name: Rafae Ash, George Chemmala} 


\vspace{.3in}
Instructions. Choose a final project and a final project partner (groups of 2 - anything different needs approval). I would like you to submit the following information:

\begin{enumerate}
	\item[(i)] \emph{Project title:} Knight's Tour
	\item[(ii)] \emph{One paragraph about the focus and scope of your project. (Don't be overly ambitious. The presentation is 10 minutes! You should pick a single question/problem/theorem to focus on.)}
		
		The presentation will focus on the Knight's tour problem and its relationship with graph theory. We will discuss how the problem of finding a knight's tour on a chessboard can be translated into a graph theory problem of finding a Hamiltonian path on the board, a subset of the Hamiltonian path problem, which in itself is a subset of the NP-hard traveling salesman problem. We will highlight the fact that while the traveling salesman problem is NP-hard, the Knight's tour problem is a special case that can be solved in linear time.

		Some topics we may briefly cover given time:
		\begin{enumerate}
			\item History
			\item Number of tours
		\end{enumerate}

	\item[(iii)] \emph{A list of at least two references you will use (not Wikipedia)}
		\begin{enumerate}
			\item[1.] \url{https://web.archive.org/web/20151010071713/http://faculty.olin.edu:80/~sadams/DM/ktpaper.pdf} (general overview)
			
			\item[2.] \url{http://www.mayhematics.com/t/t.htm} (examples/history)
			
			\item[3.] \url{https://www.tandfonline.com/doi/pdf/10.1080/0025570X.1991.11977627} (existence theorem)
			
			\item[4.] \url{https://www.sciencedirect.com/science/article/pii/0166218X9200170Q?via\%3Dihub}
		\end{enumerate}
\end{enumerate}

\end{document}
